---
title: "Negative Projection Follow-up 2"
author: "Kathryn Denning"
date: "Originally June 2019 - Updated December 2019"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
---
```{r setup data import and cleaning, echo = FALSE, warning = FALSE, message=FALSE, error=FALSE}
# Cleaning
## import data

#install.packages("rio")
#install.packages("here")
#install.packages("tidyverse")
#install.packages("magrittr")
#install.packages("janitor")
#install.packages("lme4")
#install.packages("psy")
#install.packages("irr")
#install.packages("emmeans")
#install.packages("sjPlot")
#install.packages("effects")
library(rio)
library(here)
library(tidyverse)
library(magrittr)
library(janitor)
library(lme4)
library(psy)
library(irr)
library(psych)
library(sjPlot)
library(emmeans)
library(effects)

#Setting global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

#Data after manually removing participants who did not correctly complete the PT manipulation
npf2_wide <- import("npf2_wide_ptclean2 copy.xlsx") 

#dropping second row that had full questions and overwriting because don't want that for analyses
npf2_wide <- npf2_wide[-c(1),,drop=F] 

#reverse coding
npf2_wide %<>% 
  mutate(bfi2xsh_self_quiet_r = recode(bfi2xsh_self_quiet, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_self_rude_r = recode(bfi2xsh_self_rude, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
         bfi2xsh_self_disorg_r = recode(bfi2xsh_self_disorg, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_self_taskdiff_r = recode(bfi2xsh_self_taskdiff, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_self_stable_r = recode(bfi2xsh_self_stable, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_self_abstract_r = recode(bfi2xsh_self_abstract, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
         bfi2xsh_self_risk_r = recode(bfi2xsh_self_risk, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
        bfi2xsh_self_uses_r = recode(bfi2xsh_self_uses, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_targ_quiet_r = recode(bfi2xsh_targ_quiet, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_targ_rude_r = recode(bfi2xsh_targ_rude, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
         bfi2xsh_targ_disorg_r = recode(bfi2xsh_targ_disorg, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_targ_taskdiff_r = recode(bfi2xsh_targ_taskdiff, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_targ_stable_r = recode(bfi2xsh_targ_stable, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
          bfi2xsh_targ_abstract_r = recode(bfi2xsh_targ_abstract, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
        bfi2xsh_targ_risk_r = recode(bfi2xsh_targ_risk, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1"),
        bfi2xsh_targ_uses_r = recode(bfi2xsh_targ_uses, `1` = "5",
                                   `2` = "4",
                                   `3` = "3",
                                   `4` = "2",
                                   `5` = "1")) %>% 
  select(-c(bfi2xsh_self_quiet, bfi2xsh_self_rude, bfi2xsh_self_taskdiff, 
            bfi2xsh_self_stable, bfi2xsh_self_abstract, bfi2xsh_targ_quiet, 
            bfi2xsh_targ_rude, bfi2xsh_targ_taskdiff, bfi2xsh_targ_stable, 
            bfi2xsh_targ_abstract, bfi2xsh_targ_disorg, bfi2xsh_self_disorg, 
            bfi2xsh_targ_risk, bfi2xsh_self_risk, bfi2xsh_targ_uses, 
            bfi2xsh_self_uses)) 


# putting data in long format
npf2_clean1 <- npf2_wide %>% 
  gather("bfi2xsh_type", "bfi2xsh_score", 
         c(bfi2xsh_self_heart:bfi2xsh_self_rules, 
           bfi2xsh_targ_heart:bfi2xsh_targ_rules, 
           bfi2xsh_self_quiet_r:bfi2xsh_targ_uses_r)) %>% 
  separate(.,bfi2xsh_type, c("drop",
                             "bfi2xsh_type", 
                             "bfi2xsh_qcontent"), "\\_") %>% 
  select(-drop) %>% 
  spread(bfi2xsh_type, bfi2xsh_score) %>% 
  gather("condition", "ident", c(ident_pol, 
                                 ident_race, 
                                 ident_weed, 
                                 ident_thirdpart)) %>% 
  separate(.,condition,c("drop","condition"),"\\_") %>% 
  select(-drop) %>% 
  filter(!is.na(ident)) %>% 
  gather("drop1", "lib", c(lib_pol, 
                           lib_race, 
                           lib_weed, 
                           lib_thirdpart)) %>% 
  select(-drop1) %>% 
  filter(!is.na(lib)) %>% 
  gather("drop", "cons", c(cons_pol, 
                           cons_race, 
                           cons_weed, 
                           cons_thirdpart)) %>% 
  select(-drop) %>% 
  filter(!is.na(cons)) %>% 
  gather("drop", "pt_narr", c(pt_narr_pol, 
                              pt_narr_race, 
                              pt_narr_weed, 
                              pt_narr_thirdpart)) %>% 
  select(-drop) %>% 
  #filter(!is.na(pt_narr)) %>% 
  gather("drop", "check", c(check_pol, 
                            check_race, 
                            check_weed, 
                            check_thirdpart)) %>% 
  select(-drop) %>% 
  filter(!is.na(check)) %>% 
  gather("opin_cat", "opin", c(opin_pol, 
                               opin_race, 
                               opin_weed, 
                               opin_thirdpart)) %>% 
  filter(!is.na(opin)) 


# Making the variables numeric or factor, creating perceived polarization variable, centering continous predictors, exclude people who failed manipulation or answered "neither" to politic candidates in politics condition, remove missing people with listwise missing data on variables necessary for the analysis, participants who did not complete the pt condition correctly (few word answers, said I don't know) were already removed
npf2_clean2 <- npf2_clean1 %>% 
  mutate(sub_id = as.numeric(sub_id),
         age = as.numeric(age),
         gender = as.numeric(gender),
         race = as.numeric(race),
         bfi2xsh_qcontent = as.factor(bfi2xsh_qcontent),
         self = as.numeric(self),
         targ = as.numeric(targ),
         ident = as.numeric(ident),
         lib = as.numeric(lib),
         cons = as.numeric(cons),
         condition = as.factor(condition),
         check = as.numeric(check),
         opin = as.factor(opin),
         opin_cat = as.factor(opin_cat)) %>% 
  mutate(perceived_polar = abs(lib-cons),
         condition = fct_relevel(condition, 
                                 "pol", 
                                 "race", 
                                 "weed", 
                                 "thirdpart")) %>% 
  mutate(self_c = self - mean(self, na.rm = TRUE),
         ident_c = ident - mean(ident, na.rm = TRUE),
         perceived_polar_c = perceived_polar - 
           mean(perceived_polar, na.rm = TRUE)) %>% 
  #removing opin = 5 for when including the condition variable in the analysis 
#removing it because only want in and out-group for this analysis
  filter(opin != 5) %>% 
  filter(check != 4)

```

# Demographics

## Number of participants overall

```{r participants number}
npf2_clean2 %>% 
  select(sub_id) %>% 
  unique() %>% 
  nrow()
```

## Number of participants per condition

```{r participants per condition}
npf2_clean2 %>% 
  select(sub_id, condition) %>% 
  unique() %>% 
  group_by(condition) %>% 
  count()
```

## Gender

```{r gender}
npf2_clean2 %>% 
  select(sub_id, gender) %>% 
  mutate(gender = recode(gender, 
                         `1` = "Male",
                         `2` = "Female",
                         `3` = "Other")) %>% 
  unique() %>% 
  group_by(gender) %>% 
  count()
```

## Race

```{r race}
npf2_clean2 %>% 
  select(sub_id, race) %>% 
  mutate(race = recode(race, 
                       `1` = "American Indian/Alaska Native",
                       `2` = "Asian",
                       `3` = "Black",
                       `4` = "Native Hawaiian/Pacific Islander",
                       `5` = "Caucasian",
                       `7` = "Latinx",
                       `8` = "Middle Eastern/North African",
                       `9` = "Prefer not to answer",
                       `10` = "Other")) %>% 
  unique() %>% 
  group_by(race) %>% 
  count()
```

#### Caucasian percentage

```{r}
white <- (218/336)*100
white
```

#### Asian percentage

```{r}
asian <- (47/336)*100
asian
```

#### Identify as "Other" percentage

```{r}
other <- (26/336)*100
other
```

#### Latinx percentage

```{r}
latinx <- (23/336)*100
latinx
```

#### Black percentage

```{r}
black <- (13/336)*100
black
```

All other groups did not represent more than 1% of the sample

## Age

```{r age}
npf2_clean2 %>% 
  select(sub_id, age) %>% 
  na.omit() %>% 
  summarize(mean = mean(age),
            sd = sd(age))
```

## Percent liberal

We chose to exclude people in the third party system condition because the issue is not necessarily framed as liberal or conservative issue.

```{r agree with liberal issues}
levels(npf2_clean2$opin_cat)
npf2_clean2 %>% 
  select(sub_id, opin, opin_cat) %>% 
  filter(opin_cat != "opin_thirdpart") %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(opin) %>% 
  count() 

percent_lib <- (196/238)*100
percent_lib
```

## Histogram of distributions

### In-group identification

```{r hist identification}
hist_ident <- npf2_clean2 %>% 
  select(sub_id, ident) %>% 
  unique()
hist(hist_ident$ident)
```

### Perceived polarization

```{r perceived polarization distribution}
hist_polar <- npf2_clean2 %>% 
  select(sub_id, perceived_polar) %>% 
  unique()
hist(hist_polar$perceived_polar)
```

## Mean and SD for BFI

### BFI for Self

```{r BFI self}
npf2_clean2 %>% 
  select(sub_id, condition, self) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(self),
            sd = sd(self))
```

### BFI for Target

```{r BFI targ}
npf2_clean2 %>% 
  select(sub_id, condition, targ) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(targ),
            sd = sd(targ))
```

## Mean and SD for in-group identification

```{r identification}
#Mean
npf2_clean2 %>% 
  select(sub_id, condition, ident) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(ident),
            sd = sd(ident))
```

# Model 1: Projection predicted by condition

## Contrast codes

```{r contrasts}
model_data <- npf2_clean2 %>% 
  select(targ, self_c, ident_c, condition, sub_id, perceived_polar_c, perceived_polar) %>% 
  na.omit()
levels(model_data$condition)

contrasts(model_data$condition) <- "contr.helmert"
contrasts(model_data$condition) 
```

## Results

For a result on projection, the main effect must include BFI the self (self_c) interacting with that variable.

```{r, warning = TRUE}
#Model with random slopes for BFI (aka more conservative model)
model1_randomslopes <- lmer(targ ~ self_c*ident_c*condition + (self_c|sub_id), 
                            data = model_data)

tab_model(model1_randomslopes,
          title = "MLM predicting projection with random slopes and helmert contrasts")
```

## Plots

### Main effect of in-group identification on projection

```{r model 1 plot color main effect identification}
#Getting stats for plot
describe(model_data$ident_c)

effects_data <- as_tibble(effect(term = "self_c:ident_c", 
                                 xlevels = list(ident_c = c(-0.93, 0, 0.93)),
                                 mod = model1_randomslopes), na.rm = TRUE)

effects_data$Identification <-as.factor(effects_data$ident_c)

#With confidence intervals
ggplot(effects_data, aes(self_c, fit, color = Identification)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = Identification)) +
  scale_color_manual(breaks = c(-0.93, 0, 0.93), 
                     labels = c ("Weak (1 SD Below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                     values = c('#404788FF',
                               '#55C667FF',
                               '#FDE725FF')) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 1 plot BW main effect identification}
ggplot(effects_data, aes(self_c, fit, group = Identification)) +
  geom_line(aes(linetype = Identification), size = .7) + 
  scale_linetype_manual("Identification", 
                       breaks = c(-0.93, 0, 0.93), 
                       labels = c ("Weak (1 SD below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal(base_size = 13) +
  labs(title = "Projection predicted by in-group identification",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

### Main effect of condition on projection 

```{r model 1 plot color main effect condition}
#set-up stats for plot
levels(model_data$condition) 
cond_recode_plot <- model_data %>%
  mutate(Condition = recode(condition, 
                            `pol` = "Poltics",
                            `race` = "Race",
                            `weed` = "Marijuana",
                            `thirdpart` = "Third party"))

ggplot(cond_recode_plot, aes(self_c, targ)) +
  geom_smooth(method = "lm", se = TRUE, size = 1.25, alpha = .3, aes(color = Condition, fill = Condition)) +
  theme_minimal(base_size = 13) +
  scale_color_manual(values = c('#404788FF',
                               '#238A8DFF',
                               '#55C667FF',
                               '#FDE725FF')) +
  scale_fill_manual(values = c('#404788FF',
                               '#238A8DFF',
                               '#55C667FF',
                               '#FDE725FF')) + 
  theme(legend.position = "bottom") +
  labs(title = "Main effect of issue polarization condition on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 1 plot BW main effect condition}
ggplot(cond_recode_plot, aes(self_c, targ, group = Condition)) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = .7, 
              colour = "black", 
              aes(linetype = Condition)) +
  theme_minimal(base_size = 13) +
  theme(legend.key.size = unit(1, "cm")) +
  scale_linetype_manual("Issue polarization condition",
                        labels = c("Political candidate",
                                  "Racial disadvantage", 
                                  "Marijuana legalization",
                                  "Third party system"),
                       values = c("solid",
                                   "twodash", 
                                   "dotted", 
                                   "dashed")) +
  labs(title = "Main effect of condition on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

# Model 2: Projection predicted by perceived polarization

```{r perceived pol model overall}
#model
model2_randomslopes <- lmer(targ ~ self_c*ident_c*perceived_polar_c + 
                              (self_c|sub_id), 
                            data = model_data)

#results
tab_model(model2_randomslopes, title = "Model 2 with random slopes")
```

## Plots

### Main effect of in-group identification on projection (trending)

```{r model 2 plot color main effect identification}
#Getting stats for plot
describe(model_data$ident_c)

effects_data2 <- as_tibble(effect(term = "self_c:ident_c", 
                                 xlevels = list(ident_c = c(-0.93, 0, 0.93)),
                                 mod = model2_randomslopes), na.rm = TRUE)

effects_data2$Identification <-as.factor(effects_data2$ident_c)

ggplot(effects_data2, aes(self_c, fit)) +
  geom_smooth(method = "lm", 
              size = .7,
              aes(color = Identification)) +
  scale_color_discrete(breaks = c(-0.93, 0, 0.93), 
                       labels = c ("Weak Identification \n (1 SD Below)", 
                                  "Average Identification", 
                                  "High Identification \n (1 SD above)")) +
  theme_minimal() +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 2 plot BW main effect identification}
ggplot(effects_data2, aes(self_c, fit, group = Identification)) +
  geom_line(size = .7, aes(linetype = Identification)) + 
  scale_linetype_manual("Identification", 
                       breaks = c(-0.93, 0, 0.93), 
                       labels = c ("Weak Identification \n (1 SD Below)", 
                                  "Average Identification", 
                                  "High Identification \n (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal() +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

### Main effect of perceived polarization on projection (trending)

```{r model 2 plot color main effect perceived polar}
#Getting stats for plot
describe(model_data$perceived_polar)

effects_data3 <- as_tibble(effect(term = "self_c:perceived_polar_c", 
                                 xlevels = list(perceived_polar_c = c(-1.96, 
                                                                      0, 
                                                                      1.96)),
                                 mod = model2_randomslopes), na.rm = TRUE)

effects_data3$perceived_polar <-as.factor(effects_data3$perceived_polar_c)

ggplot(effects_data3, aes(self_c, fit)) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = 1.25, 
              aes(color = perceived_polar)) +
  scale_color_viridis_d("Perceived Polarization",
                       breaks = c(-1.96, 0, 1.96), 
                       labels = c ("Weak (1 SD Below)", 
                                  "Average", 
                                  "High (1 SD above)")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") +
  labs(title = "Trending main effect of perceived polarization on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 2 plot BW main effect perceived polar}
ggplot(effects_data3, aes(self_c, fit)) +
  geom_smooth(method = "lm", 
              size = .7, 
              colour = "black",  
              aes(linetype = perceived_polar)) +
  scale_linetype_manual("Perceived Polarization",
                        breaks = c(-1.96, 0, 1.96), 
                       labels = c ("Low (1 SD below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal() +
  labs(title = "Main effect of perceived polarization predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

# Perceived polarization as outcome predicted by condition

## Descriptives

```{r perceived polarization by condition without neither}
mean_perceivedpol <- model_data %>% 
  select(sub_id, condition, perceived_polar) %>% 
   mutate(condition = recode(condition, 
                            `pol` = "Poltical candidate",
                            `race` = "Racial disadvantage",
                            `weed` = "Marijuana legalization",
                            `thirdpart` = "Third party system")) %>% 
  group_by(condition) %>%
  mutate(mean_percbycond = mean(perceived_polar)) %>% 
  summarise(mean = mean(mean_percbycond),
            sd = sd(perceived_polar))
mean_perceivedpol
```

### Results

```{r perceived pol outcome model}
#contrast codes
contrasts(npf2_clean2$condition) <- "contr.helmert"
contrasts(npf2_clean2$condition) 

polar_outcome <- lm(perceived_polar ~ condition, data = npf2_clean2)
summary(polar_outcome)
confint(polar_outcome)
```

```{r plots perceived pol outcome model}
ggplot(mean_perceivedpol, aes(x = condition, y = mean)) +
  geom_bar(stat="identity", aes(color = condition)) +
  theme(legend.position = "none") +
  scale_colour_grey() +
  theme_minimal(base_size = 13) +
  labs(title = "Percieved polarization predicted by condition",
       x = "Issue polarization condition",
       y = "Perceived polarization") +
  theme(legend.position="none")
```
