---
title: "Negative Projection Follow-up 2"
author: "Kathryn Denning"
date: "Originally June 2019 - Updated December 2019"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
---
```{r setup data import and cleaning, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
# Cleaning
## import data

#install.packages("rio")
#install.packages("here")
#install.packages("tidyverse")
#install.packages("magrittr")
#install.packages("janitor")
#install.packages("lme4")
#install.packages("psy")
#install.packages("irr")
#install.packages("emmeans")
#install.packages("sjPlot")
#install.packages("effects")
library(rio)
library(here)
library(tidyverse)
library(magrittr)
library(janitor)
library(lme4)
library(psy)
library(irr)
library(psych)
library(sjPlot)
library(emmeans)
library(effects)

# Setting global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

# Data after manually removing participants who did not correctly complete the PT manipulation
npf2_wide <- import("npf2_wide_ptclean2 copy.xlsx") 

# Dropping second row that had full questions and overwriting because don't want that for analyses
npf2_wide <- npf2_wide[-c(1),, drop = F] 

# Reverse coding
npf2_wide_recode <- npf2_wide %>% 
     mutate_at(c("bfi2xsh_self_quiet", "bfi2xsh_self_rude", 
                 "bfi2xsh_self_disorg", "bfi2xsh_self_taskdiff",
                 "bfi2xsh_self_stable", "bfi2xsh_self_abstract",
                 "bfi2xsh_self_risk", "bfi2xsh_self_uses",
                 "bfi2xsh_targ_quiet", "bfi2xsh_targ_rude",
                 "bfi2xsh_targ_rude", "bfi2xsh_targ_disorg",
                 "bfi2xsh_targ_taskdiff", "bfi2xsh_targ_stable",
                 "bfi2xsh_targ_abstract", "bfi2xsh_targ_risk",
                 "bfi2xsh_targ_uses"), 
               list(~recode(., `1`= 5, `2`= 4, `3` = 3, `4` = 2, `5` = 1)))

# Cleaning and putting data in long-format
npf2_clean <- npf2_wide_recode %>% 
  # Dropping open-ended questions
  select(-c(pt_narr_pol, pt_narr_race, pt_narr_thirdpart, pt_narr_weed)) %>% 
  # Making all character variables numeric
  mutate_if(is.character,as.numeric) %>% 
  # Putting responses for personality items into long format for participant and target
  pivot_longer(c(bfi2xsh_self_quiet:bfi2xsh_self_rules, 
           bfi2xsh_targ_quiet:bfi2xsh_targ_rules),
           names_sep = "_",
           names_to = c("drop1", "bfi2xsh_type", "bfi2xsh_qtype")) %>% 
  pivot_wider(names_from = bfi2xsh_type, values_from = value) %>% 
  # Putting responses on other items (conditon, identification, liberal and conservative ratings) into long-format
  pivot_longer(c(opin_pol:check_thirdpart),
               names_sep = "_",
               names_to = c("question", "condition")) %>% 
  na.omit() %>% 
  pivot_wider(names_from = question, values_from = value) %>% 
  select(-drop1) %>% 
  # Making a varriable for perceived polarization, making condition a factor variable
  mutate(perceived_polar = abs(lib-cons),
         condition = fct_relevel(condition, 
                                 "pol", 
                                 "race", 
                                 "weed", 
                                 "thirdpart")) %>% 
  # Mean centering continuous predictors
  mutate(self_c = self - mean(self, na.rm = TRUE),
         ident_c = ident - mean(ident, na.rm = TRUE),
         perceived_polar_c = perceived_polar - 
           mean(perceived_polar, na.rm = TRUE)) %>% 
# Removing opin = 5 for when including the condition variable in the analysis 
# Removing it because only want in and out-group for this analysis, this was an error in data collection to have a third option
  filter(opin != 5) %>% 
  # Removing those who responded incorrectly to the manipulation check
  filter(check != 4)
```

# Demographics

## Number of participants overall

```{r participants number}
npf2_clean %>% 
  select(sub_id) %>% 
  unique() %>% 
  nrow()
```

## Number of participants per condition

```{r participants per condition}
npf2_clean %>% 
  select(sub_id, condition) %>% 
  unique() %>% 
  group_by(condition) %>% 
  count()
```

## Gender

```{r gender}
npf2_clean %>% 
  select(sub_id, gender) %>% 
  mutate(gender = recode(gender, 
                         `1` = "Male",
                         `2` = "Female",
                         `3` = "Other")) %>% 
  unique() %>% 
  group_by(gender) %>% 
  count()
```

## Race

```{r race}
npf2_clean %>% 
  select(sub_id, race) %>% 
  mutate(race = recode(race, 
                       `1` = "American Indian/Alaska Native",
                       `2` = "Asian",
                       `3` = "Black",
                       `4` = "Native Hawaiian/Pacific Islander",
                       `5` = "Caucasian",
                       `7` = "Latinx",
                       `8` = "Middle Eastern/North African",
                       `9` = "Prefer not to answer",
                       `10` = "Other")) %>% 
  unique() %>% 
  group_by(race) %>% 
  count() %>% 
  mutate(percent = n/334*100)
```

## Age

```{r age}
npf2_clean %>% 
  select(sub_id, age) %>% 
  na.omit() %>% 
  summarize(mean = mean(age),
            sd = sd(age))
```

## Percent liberal

We chose to exclude people in the third party system condition because the issue is not necessarily framed as liberal or conservative issue.

```{r agree with liberal issues}
npf2_clean %>% 
  select(sub_id, opin, condition) %>% 
  filter(condition != "thirdpart") %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(opin) %>% 
  count() %>% 
  mutate(percent_lib = n/236*100)
```

## Histogram of distributions

### In-group identification

```{r hist identification}
hist_ident <- npf2_clean %>% 
  select(sub_id, ident) %>% 
  unique()
hist(hist_ident$ident)
```

### Perceived polarization

```{r perceived polarization distribution}
hist_polar <- npf2_clean %>% 
  select(sub_id, perceived_polar) %>% 
  unique()
hist(hist_polar$perceived_polar)
```

## Mean and SD for BFI

### BFI for Self

```{r BFI self}
npf2_clean %>% 
  select(sub_id, condition, self) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(self),
            sd = sd(self))
```

### BFI for Target

```{r BFI targ}
npf2_clean %>% 
  select(sub_id, condition, targ) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(targ),
            sd = sd(targ))
```

## Mean and SD for in-group identification

```{r identification}
npf2_clean %>% 
  select(sub_id, condition, ident) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(ident),
            sd = sd(ident))
```

# Model 1: Projection predicted by condition

## Contrast codes

```{r contrasts}
model_data <- npf2_clean %>% 
  select(targ, self_c, ident_c, condition, sub_id, perceived_polar_c, perceived_polar) %>% 
  na.omit()
levels(model_data$condition)

contrasts(model_data$condition) <- "contr.helmert"
contrasts(model_data$condition) 
```

## Results

For a result on projection, the main effect must include BFI the self (self_c) interacting with that variable.

```{r, warning = TRUE}
#Model with random slopes for BFI (aka more conservative model)
model1_randomslopes <- lmer(targ ~ self_c*ident_c*condition + (self_c|sub_id), 
                            data = model_data)

tab_model(model1_randomslopes,
          title = "MLM predicting projection with random slopes and helmert contrasts",
          dv.labels = "BFI responses for target",
          pred.labels = c("(Intercept)",
                          "BFI responses for self",
                          "Identification",
                          "Condition 1: Politics vs Race",
                          "Condition 2: Politics, Race vs Marijuana",
                          "Condition 3: Politics, Race, Marijuana vs Third Party",
                          "BFI responses for self:Identification",
                          "BFI responses for self:Condition 1",
                          "BFI responses for self:Condition 2",
                          "BFI responses for self:Condition 3",
                          "Identification:Conditon 1",
                          "Identification:Condition 2",
                          "Identification:Condition 3",
                          "BFI responses for self:Identification:Condition 1",
                          "BFI responses for self:Identification:Condition 2",
                          "BFI responses for self:Identification:Condition 3"))
```

## Plots for Model 1

### Main effect of in-group identification on projection

```{r model 1 plot color main effect identification}
#Getting stats for plot
describe(model_data$ident_c)

effects_data <- as_tibble(effect(term = "self_c:ident_c", 
                                 xlevels = list(ident_c = c(-0.94, 0, 0.94)),
                                 mod = model1_randomslopes), na.rm = TRUE)

effects_data$Identification <-as.factor(effects_data$ident_c)

# Color plot for conference presentation
ggplot(effects_data, aes(self_c, fit, color = Identification)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = Identification)) +
  scale_color_manual(breaks = c(-0.94, 0, 0.94), 
                     labels = c ("Weak (1 SD Below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                     values = c('#404788FF',
                               '#55C667FF',
                               '#FDE725FF')) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 1 plot BW main effect identification}
# Black and white plot for publication
ggplot(effects_data, aes(self_c, fit, group = Identification)) +
  geom_line(aes(linetype = Identification), size = .7) + 
  scale_linetype_manual("Identification", 
                       breaks = c(-0.94, 0, 0.94), 
                       labels = c ("Weak (1 SD below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal(base_size = 13) +
  labs(title = "Projection predicted by in-group identification",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

### Main effect of condition on projection 

```{r model 1 plot color main effect condition}
# Set-up stats for plot
levels(model_data$condition) 
cond_recode_plot <- model_data %>%
  mutate(Condition = recode(condition, 
                            `pol` = "Poltics",
                            `race` = "Race",
                            `weed` = "Marijuana",
                            `thirdpart` = "Third party"))

# Color plot for conference presentation
ggplot(cond_recode_plot, aes(self_c, targ)) +
  geom_smooth(method = "lm", se = TRUE, size = 1.25, alpha = .3, aes(color = Condition, fill = Condition)) +
  theme_minimal(base_size = 13) +
  scale_color_manual(values = c('#404788FF',
                               '#238A8DFF',
                               '#55C667FF',
                               '#FDE725FF')) +
  scale_fill_manual(values = c('#404788FF',
                               '#238A8DFF',
                               '#55C667FF',
                               '#FDE725FF')) +
  labs(title = "Main effect of issue polarization condition on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 1 plot BW main effect condition}
# Black and white plot for publication
ggplot(cond_recode_plot, aes(self_c, targ, group = Condition)) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = .7, 
              colour = "black", 
              aes(linetype = Condition)) +
  theme_minimal(base_size = 13) +
  theme(legend.key.size = unit(1, "cm")) +
  scale_linetype_manual("Issue polarization",
                        labels = c("Political candidate",
                                  "Racial disadvantage", 
                                  "Marijuana legalization",
                                  "Third party system"),
                       values = c("solid",
                                   "twodash", 
                                   "dotted", 
                                   "dashed")) +
  labs(title = "Main effect of issue polarization on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

# Model 2: Projection predicted by perceived polarization

```{r perceived pol model overall}
# Model
model2_randomslopes <- lmer(targ ~ self_c*ident_c*perceived_polar_c + 
                              (self_c|sub_id), 
                            data = model_data)

# Results
tab_model(model2_randomslopes, title = "Model 2 with random slopes",
          dv.labels = "BFI responses for target",
          pred.labels = c("(Intercept)",
                          "BFI for self",
                          "Identification",
                          "Perceived polarization",
                          "BFI for self:Identification",
                          "BFI for self:Perceived polarization",
                          "Identification:Perceived polarization",
                          "BFI for self:Identification:Perceived polarization"))
```

## Plots for Model 2

### Main effect of in-group identification on projection (trending)

```{r model 2 plot color main effect identification}
#Getting stats for plot
describe(model_data$ident_c)

effects_data2 <- as_tibble(effect(term = "self_c:ident_c", 
                                 xlevels = list(ident_c = c(-0.94, 0, 0.94)),
                                 mod = model2_randomslopes), na.rm = TRUE)

effects_data2$Identification <-as.factor(effects_data2$ident_c)

# Color plot
ggplot(effects_data2, aes(self_c, fit)) +
  geom_line(size = .7, aes(color = Identification)) +
  scale_color_viridis_d(breaks = c(-0.94, 0, 0.94), 
                       labels = c ("Weak Identification \n (1 SD Below)", 
                                  "Average Identification", 
                                  "High Identification \n (1 SD above)")) +
  theme_minimal() +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 2 plot BW main effect identification}
ggplot(effects_data2, aes(self_c, fit, group = Identification)) +
  geom_line(size = .7, aes(linetype = Identification)) + 
  scale_linetype_manual("Identification", 
                       breaks = c(-0.94, 0, 0.94), 
                       labels = c ("Weak Identification \n (1 SD Below)", 
                                  "Average Identification", 
                                  "High Identification \n (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal() +
  labs(title = "Main effect of in-group identification predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

### Main effect of perceived polarization on projection (trending)

```{r model 2 plot color main effect perceived polar}
# Getting stats for plot
describe(model_data$perceived_polar)

effects_data3 <- as_tibble(effect(term = "self_c:perceived_polar_c", 
                                 xlevels = list(perceived_polar_c = c(-1.96, 
                                                                      0, 
                                                                      1.96)),
                                 mod = model2_randomslopes), na.rm = TRUE)

effects_data3$perceived_polar <-as.factor(effects_data3$perceived_polar_c)

# Color plot for conference presentation
ggplot(effects_data3, aes(self_c, fit)) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              size = 1.25, 
              aes(color = perceived_polar)) +
  scale_color_viridis_d("Perceived Polarization",
                       breaks = c(-1.96, 0, 1.96), 
                       labels = c ("Weak (1 SD Below)", 
                                  "Average", 
                                  "High (1 SD above)")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom") +
  labs(title = "Trending main effect of perceived polarization on projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

```{r model 2 plot BW main effect perceived polar}
# Black and white plot for publication
ggplot(effects_data3, aes(self_c, fit)) +
  geom_line(size = .7, 
              colour = "black",  
              aes(linetype = perceived_polar)) +
  scale_linetype_manual("Perceived Polarization",
                        breaks = c(-1.96, 0, 1.96), 
                       labels = c ("Low (1 SD below)", 
                                  "Average", 
                                  "High (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal() +
  labs(title = "Main effect of perceived polarization predicting projection",
       x = "Personality ratings for self",
       y = "Personality ratings for target")
```

# Perceived polarization as outcome predicted by condition

## Descriptives

```{r perceived polarization by condition without neither}
mean_perceivedpol <- model_data %>% 
  select(sub_id, condition, perceived_polar) %>% 
   mutate(condition = recode(condition, 
                            `pol` = "Poltical candidate",
                            `race` = "Racial disadvantage",
                            `weed` = "Marijuana legalization",
                            `thirdpart` = "Third party system")) %>% 
  group_by(condition) %>%
  mutate(mean_percbycond = mean(perceived_polar)) %>% 
  summarise(mean = mean(mean_percbycond),
            sd = sd(perceived_polar))
mean_perceivedpol
```

## Results

```{r perceived pol outcome model}
# Contrast codes
contrasts(npf2_clean$condition) <- "contr.helmert"
contrasts(npf2_clean$condition) 

# Model
polar_outcome <- lm(perceived_polar ~ condition, data = npf2_clean)
summary(polar_outcome)
confint(polar_outcome)
```

```{r plots perceived pol outcome model}
ggplot(mean_perceivedpol, aes(x = condition, y = mean)) +
  geom_bar(stat = "identity") +
  theme(legend.position = "none") +
  scale_colour_grey() +
  theme_minimal(base_size = 13) +
  labs(title = "Percieved polarization predicted by condition",
       x = "Issue polarization condition",
       y = "Perceived polarization") +
  theme(legend.position = "none")
```
